Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-31 transmission-appendix-INLA.tex
proof

An LGM is a model for a series of observations $\vec{y} = [y_1, \dots, y_n]^T$.
Each observation has an associated linear predictor $\vec{\eta} = [\eta_1, \dots, \eta_n]^T$.
The linear predictor and, optionally, some hyperparameters $\vec{\theta}$ give the conditional independence structure:
\begin{align}
    \pi(\vec{y} \mid \vec{\eta}, \vec{\theta}) = \prod_{i=1}^n \pi(y_i \mid \eta_i, \vec{\theta}).
\end{align}
Such a structure is familiar from linear regression models which have $\vec{\eta} = \matr{X} \vec{\beta}$, where $\matr{X}$ is the design matrix and $\vec{\beta}$ are the coefficients; $\vec{\theta}$ is one-dimensional; and $\pi(y_i \mid \eta_i, \vec{\theta})$ is the normal distribution with mean $\eta_i$ and standard deviation $\theta$.

An LGM introduces a latent field $\vec{x}$ with a multivariate Gaussian distribution (hence the name):
\begin{align}
    \vec{x} \mid \vec{\theta} \dist \N(\vec{0}, \matr{\Sigma}(\vec{\theta}))
\end{align}
where $\Sigma(\vec{\theta})$ is the covariance matrix formed as a deterministic transformation of the hyperparameters.
The linear predictors must be able to be expressed as a linear transformation of the latent field, \ie $\vec{\eta} = \matr{A} \vec{x}$ for some pre-specified matrix $\matr{A}$.
For the linear regression model, $\matr{A} = \matr{X}$, $\vec{x} = \vec{\beta}$, and $\matr{\Sigma}(\vec{\theta})$ would be a fixed diagonal matrix specifying the prior on the regression coefficients.
In the (generalised) linear model case $\vec{\theta}$ is fixed; however, LGMs allow an arbitrary prior distribution on $\vec{\theta}$ which means that some elements of $\vec{x}$ can be random effects (or more complex structures).

>>>>

An LGM is a model for a series of observations $\vec{y} = [y_1, \dots, y_n]^T$.
Each observation has an associated linear predictor $\vec{\eta} = [\eta_1, \dots, \eta_n]^T$.
The linear predictor and, optionally, some hyperparameters $\vec{\theta}$ have the conditional independence structure:
\begin{align}
    \pi(\vec{y} \mid \vec{\eta}, \vec{\theta}) = \prod_{i=1}^n \pi(y_i \mid \eta_i, \vec{\theta}).
\end{align}
Such a structure is familiar from linear regression models which have $\vec{\eta} = \matr{X} \vec{\beta}$, where $\matr{X}$ is the design matrix and $\vec{\beta}$ are the coefficients; $\vec{\theta}$ is one-dimensional; and $\pi(y_i \mid \eta_i, \vec{\theta})$ is the normal distribution with mean $\eta_i$ and standard deviation $\theta$.

An LGM introduces a latent field $\vec{x}$ with a multivariate Gaussian distribution (hence the name):
\begin{align}
    \vec{x} \mid \vec{\theta} \dist \N(\vec{0}, \matr{\Sigma}(\vec{\theta}))
\end{align}
where $\Sigma(\vec{\theta})$ is the covariance matrix formed as a deterministic transformation of the hyperparameters.
The linear predictors must be able to be expressed as a linear transformation of the latent field, \ie $\vec{\eta} = \matr{A} \vec{x}$ for some pre-specified matrix $\matr{A}$.
For the linear regression model, $\matr{A} = \matr{X}$, $\vec{x} = \vec{\beta}$, and $\matr{\Sigma}(\vec{\theta})$ would be a fixed diagonal matrix specifying the prior on the regression coefficients.
In the (generalised) linear model case $\matr{\Sigma(\vec{\theta})}$ is fixed; however, LGMs allow an arbitrary prior distribution on $\vec{\theta}$ which means that some elements of $\vec{x}$ can be random effects (or more complex structures).

### 2024-03-31 transmission.tex
proof

Properties 3 and 5 suggest the use of an approximate method for computing the posterior distribution.
A particular issue is that using MCMC, sufficient covariates to adjust for the non-response bias, and a period of more than 6--8 weeks leads to a model that is computationally prohibitive ~\citePersonalComms{Koen Pouwels}.
Instead, previous MRP implementations for the CIS used INLA (see \cref{E-biology-data:sec:MRP}).

INLA-based inference can only be applied to LGMs (\emph{Latent Gaussian Models}); see \cref{E-transmission:sec:INLA} for further details on INLA and LGMs.
The LGM framework requires that a transformation of $P_{i,t}$ is linear in $\vec{g} = [g_{{\vec{\varphi}}}(1, 1), \dots, g_{{\vec{\varphi}}}(1, T), \dots, g_{{\vec{\varphi}}}(2, 1), \dots, g_{{\vec{\varphi}}}(I, T)$ where $I$ and $T$ are the number of strata and days respectively.

@@

Then, I transform this posterior sample using \cref{transmission:alg:calculate-Z}.
The algorithm is based on the forward-substitution algorithm from \cref{inc-prev:eq:forward-substitute}, which including stratum is:
\begin{align}
Z_{i,t}
&= P_{i,t} - \sum_{j=1}^{t-1} S(t - j + 1) Z_{i,j}.
\label{transmission:eq:forward-substitute}
\end{align}
Dividing each side of \cref{inc-prev:eq:forward-substitute} by the total number of individuals in each stratum transforms the number of prevalent individuals to prevalence and incidence to incidence proportion.
The algorithm uses a deterministic relationship between incidence and prevalence within a stratum, as \cref{E-inc-prev:sec:observation-process} justifies; \ie, the estimate of the incidence is negligibly affected by approximating the distribution $\vec{P} \mid \vec{Z}$ by its mean.

\begin{algorithm}
    \For{$j = 1, \dots, 500$}{
      Set $\vec{P}^{(j)}$ to a posterior sample of $\vec{P}$ \;
      \For{each stratum $i$}{
        \For{$t = -\dmax, \dots, 0$}{
            $Z^{(j)}_{i,t} = P^{(j)}_{i,1} / \sum_{k=1}^{\dmax} S(k)$
        }
        \For{$t = 1, \dots, T$}{
            Set $Z^{(j)}_{i,t} = P^{(j)}_{i,t} - \sum_{k=-\dmax}^{t-1} S(t - k + 1) Z^{(j)}_{i,k}$ \;
        }
      }
    }
    \caption{%
      Algorithm to transform 500 posterior samples posterior samples of $\vec{P}$ to posterior samples of $\vec{Z}$.
      Assumes constant incidence for $t \leq 1$ where $t = 1$ is the first day of prevalence data.
      Calculation of $Z^{(j)}_{i,t}$ is from \cref{transmission:eq:forward-substitute}.
    }
    \label{transmission:alg:calculate-Z}
\end{algorithm}

@@

Violation of assumptions in various parts of this procedure can lead to the second term in \cref{transmission:eq:forward-substitute} to be greater than the first, giving negative incidence.
In this case, the incidence prior to day $t$ imply that the prevalence on day $t$ is higher than that predicted by the fitted regression model for this strata/day combination.

@@

The incidence proportion was highest among the 12--16 and 17--25 age groups throughout the period (see \cref{transmission:fig:backcalc-ages}).
Except the 17--25 age group, all age groups show similar trends (\ie parallel lines on a log scale), meaning there was insufficient information in the data to overcome the prior which favoured this situation.
The incidence proportion in the 17--25 age group decreases faster than other age groups during November.
There are many plausible explanations for this effect.
The higher incidence in this age group prior to November means it could be an immunity effect.
Alternatively, it could be a consequence of schools remaining open in the November lockdown; 20--24 year olds rarely interact with school-aged children (see \cref{E-SEIR:fig:age-contacts})
Many other explanations are possible, and further work is required to determine causality.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{transmission/backcalc-ages}}
    \caption[Incidence estimated using the phenomenological approach by age group]{%
        Incidence proportion for each age group, on both natural and log scales.
    }
    \label{transmission:fig:backcalc-ages}
\end{figure}

An edge effect, a rapid rise in incidence between days one and two, is visible (see \cref{transmission:fig:backcalc-start-effect}).
This is due to the incorrect assumption that incidence is constant in the past.
Changing this assumption (\eg to exponential growth) modifies the estimated value on the first day, but the rest of the series is insensitive to this change (not shown).
Therefore, the first day's estimate is unreliable, and has been excluded from all other results.
However, the rest of the series is robust to this assumption.
\begin{figure}
    \centering \includegraphics{transmission/backcalc-start-effect}
    \caption[Edge effects in the phenomenological approach method]{%
        First seven days of incidence proportion estimate for the whole of England, highlighting the edge effect.
    }
    \label{transmission:fig:backcalc-start-effect}
\end{figure}

@@

In this section, I apply the general mechanistic transmission modelling framework from \cref{E-SEIR} to the CIS prevalence data.
The application of this framework follows \textcite{birrellRealtime}.
The largest modification from the generic framework presented in \cref{E-SEIR} is the parameterization of the WAIFW matrix (\cref{SEIR:sec:WAIFW-application}).
A novel approach (\cref{SEIR:sec:observation}) links this transmission model to the CIS prevalence data; the approach is generalisable to other studies collecting representative prevalence data, \eg REACT.
In the mechanistic modelling literature, this part of the model is referred to as the observation model, however, in the terminology of \cref{E-inc-prev}, it includes both the prevalence process and the observation process.
Having specified the full model (\cref{SEIR:sec:methods-application}), I conducted a simulation study to validate the inference process and identifiability (\cref{SEIR:sec:sim-study}), before applying the model to the CIS data (\cref{SEIR:sec:application}).
In \cref{SEIR:sec:discussion}, I discuss the results; in \cref{transmission:sec:comparison} I compare the results to those in the previous section.

@@

I link the modelled proportion of individuals that are RT-PCR positive to the data using a beta-binomial likelihood (as discussed in \cref{E-biology-data:sec:clustering}).
If on each day $t$, each stratum $i$ is observed to have $y_{at}$ positives out of $n_{at}$ tests, then the likelihood is:
\begin{align}
    \prod_{a,t} \BB (y_{at} \mid n_{at}, \bar{p}_{at}, \rho)
    \label{SEIR:eq:likelihood}
\end{align}
where: $\bar{p}_{at}$ is the mean proportion of individuals in strata $a$ that are RT-PCR positive on day $t$; and $\rho$ is a nuisance parameter controlling the overdispersion of the beta-binomial distribution.
Denoting the proportion of stratum $a$ in state $p_k$ as $p_{ka}(\tau)$, then $\bar{p}_{at} = \int_{t}^{t+1} \sum_{k=1}^l p_{ka}(\tau) d\tau$.

@@

Equating the expressions in \cref{SEIR:eq:fullODEs} and \cref{SEIR:eq:exponential-growth}, gives the following solution for the initial state of several compartments.
\begin{align}
    \vec{e}_2(0) &= \vec{i}_1(0) \frac{\psi + 2\gamma}{2\sigma} \\
    \vec{e}_1(0) &= \vec{e}_2(0) \frac{\psi + 2\sigma}{2\sigma} \\
    \vec{p}_0(0) &= \vec{e}_1(0) \frac{\psi + 2\sigma}{\psi + \omega_0} \\
    \vec{p}_1(0) &= \vec{p}_0(0) \frac{\omega_0}{\psi + \omega_1} \\
    \vec{p}_2(0) &= \vec{p}_1(0) \frac{\omega_1}{\psi + \omega_{2}} \\
    \vec{p}_3(0) &= \vec{p}_2(0) \frac{\pi \omega_{2}}{\psi + \omega_{3}}.
\end{align}

@@

I compute the posterior distribution using MCMC.
Specifically, I use an adaptive random-walk Metropolis--Hastings algorithm (discussed in \cref{E-MCMC}).
The proposal distribution is a multivariate normal that adapts to the covariance of the posterior.
The adaptive algorithm is based on \textcite[algorithm 4]{andrieuTutorial}, with the implementation based on \textcite{ghoshApproximate}'s work.
Proposals are accepted or rejected using the standard Metropolis--Hastings (MH) acceptance probability.
See \cref{SEIR:MCMC-algorithm} for full details.

@@

I ran a simulation study to study the inference procedure's performance and the model's practical identifiability.
I generated 100 simulated datasets representing the East of England using the following procedure.

@@

Next, I solved the ODE system in \cref{SEIR:eq:fullODEs} for each parameter set.
All other aspects of the ODEs, including the data-informed contact matrices, $\matr{\kappa}$, are as previously described in this section.
Population sizes were those of the East of England, and the time period the same as the main analysis.
This means that the simulations created realistic incidence and prevalence curves (see \cref{SEIR:fig:sim-data}).
In particular, the October half-term school holiday is visible, especially in the school-age groups and the November lockdown is visible to some extent in all age groups.
\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/sim/data}}
    \caption[Simulated data]{%
        (A) Incidence and (B) prevalence for the 100 simulated data sets as a proportion of each stratum.
        Prevalence shown is $\bar{p}_{kat}$, the true population prevalence on each day, excluding sampling noise.
        Each line represents one simulation.
    }
    \label{SEIR:fig:sim-data}
\end{figure}

@@

Having generated 100 datasets, I used the inference procedure in \cref{SEIR:sec:inference} to find the posterior distribution conditional on each dataset.
Hence, I ended up with 100 parameter sets distributed according to the priors (the ``generative values"), with an associated dataset and posterior distribution for it.

@@

In this set-up, the parameters $i^+$, $q_c$, and $\psi$ are fully identified; $\pi_0$ is not identified at all; the other parameters are partially identified, their posterior has moved away from the prior but there is insufficient information in the data to fully overcome the prior (see \cref{SEIR:fig:true-vs-posterior}).
The ability to identify $q_c$ is an advantage of using data which includes measurements on children, unlike data on hospitalisations or deaths.

@@

The \emph{attack rate} is the proportion of the population that have been infected at least once.
In this model it is $1 - \vec{s}$, the proportion susceptible at the time the study ends.
The attack rate is highest in the 12--16 age group in the North West at 68\% (95\% CrI: 66--70\%) (see \cref{SEIR:fig:attack-rates}).
Some of these infections occurred prior to the period covered by this study; during the period of this study 23\% (95\% CrI: 21--24\%) of the population in the North West aged 11--16 was infected.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{SEIR/CIS/attack_rates}}
    \caption{Posterior median and 95\% CrI of the attack rate in each region and age group.}
    \todo[inline]{Maybe compare this to blood donors, as used in the prior.}
    \label{SEIR:fig:attack-rates}
\end{figure}

@@

These patterns can be interpreted as an instance of the common bias---variance trade-off.
The SEIR model imposes stronger assumptions about the epidemic dynamics.
The assumptions reduce the variance of the estimates, including by making them smoother, but if the assumptions are incorrect, bias is introduced.
Furthermore, the SEIR model does not fully adjust for the non-response bias.
In particular, it makes no adjustment for the under-representation of ethnic minority groups in the CIS.
Adjusting for more covariates would require further stratifying the population and finding an appropriate parameterization of the contact matrix.
This trade-off based on the strength of assumptions has long been known to apply in the backcalculation context~\autocite[e.g.][section 8.3]{brookmeyerBackcalculation}.
The SEIR model can be viewed as a backcalculation method where the transmission part of the model is a smoothing method.

>>>>

Properties 3 and 5 suggest the use of an approximate method for computing the posterior distribution.
A particular issue is that using MCMC, sufficient covariates to adjust for the non-response bias, and a period of more than 6--8 weeks leads to a model that is computationally prohibitive ~\citePersonalComms{Koen Pouwels}.
Instead, previous MRP implementations for the CIS used INLA (as in \cref{E-biology-data:sec:MRP}).

INLA-based inference can only be applied to LGMs (\emph{Latent Gaussian Models}); see \cref{E-transmission:sec:INLA} for further details on INLA and LGMs.
The LGM framework requires that a transformation of $P_{i,t}$ is linear in $\vec{g} = [g_{{\vec{\varphi}}}(1, 1), \dots, g_{{\vec{\varphi}}}(1, T), \dots, g_{{\vec{\varphi}}}(2, 1), \dots, g_{{\vec{\varphi}}}(I, T)]^T$ where $I$ and $T$ are the number of strata and days respectively.

@@

Then, I transform this posterior sample using \cref{transmission:alg:calculate-Z}.
The algorithm is based on the forward-substitution algorithm from \cref{inc-prev:eq:forward-substitute}, which including stratum is:
\begin{align}
Z_{i,t}
&= P_{i,t} - \sum_{j=1}^{t-1} S(t - j + 1) Z_{i,j}.
\label{transmission:eq:forward-substitute}
\end{align}
Dividing each side of \cref{inc-prev:eq:forward-substitute} by the total number of individuals in each stratum transforms the number of prevalent individuals to prevalence and incidence to incidence proportion.
The algorithm uses a mean-field approximation to the relationship between incidence and prevalence within a stratum, as \cref{E-inc-prev:sec:observation-process} justifies.

\begin{algorithm}
    \For{$j = 1, \dots, 500$}{
      Set $\vec{P}^{(j)}$ to a posterior sample of $\vec{P}$ \;
      \For{each stratum $i$}{
        \For{$t = -\dmax, \dots, 0$}{
            Set $Z^{(j)}_{i,t} = P^{(j)}_{i,1} / \sum_{k=1}^{\dmax} S(k)$ \;
        }
        \For{$t = 1, \dots, T$}{
            Set $Z^{(j)}_{i,t} = P^{(j)}_{i,t} - \sum_{k=-\dmax}^{t-1} S(t - k + 1) Z^{(j)}_{i,k}$ \;
        }
      }
    }
    \caption{%
      Algorithm to transform 500 posterior samples posterior samples of $\vec{P}$ to posterior samples of $\vec{Z}$.
      Assumes constant incidence for $t \leq 1$ where $t = 1$ is the first day of prevalence data.
      Calculation of $Z^{(j)}_{i,t}$ is from \cref{transmission:eq:forward-substitute}.
    }
    \label{transmission:alg:calculate-Z}
\end{algorithm}

@@

Violation of assumptions in various parts of this procedure can lead to the second term in \cref{transmission:eq:forward-substitute} to be greater than the first, giving negative incidence.
In this case, the incidence prior to day $t$ imply that the prevalence on day $t$ is higher than that predicted by the fitted regression model for this strata/day combination.
In this case, I replace the negative value for $Z_{i,t}$ with zero.

@@

The incidence proportion was highest among the 12--16 and 17--25 age groups throughout the period (see \cref{transmission:fig:backcalc-ages}).
Except the 17--25 age group, all age groups show similar trends (\ie parallel lines on a log scale), meaning there was insufficient information in the data to overcome the prior which favoured this situation.
The incidence proportion in the 17--25 age group decreases faster than other age groups during November.
There are many plausible explanations for this effect.
The higher incidence in this age group prior to November means it could be an immunity effect.
Alternatively, it could be a consequence of schools remaining open in the November lockdown; 20--24 year olds rarely interact with school-aged children (see \cref{E-SEIR:fig:age-contacts})
Further work is required to determine causality.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{transmission/backcalc-ages}}
    \caption[Incidence estimated using the phenomenological approach by age group]{%
        Incidence proportion for each age group, on both natural and log scales.
    }
    \label{transmission:fig:backcalc-ages}
\end{figure}

An edge effect, a rapid rise in incidence between days one and two, is visible (see \cref{transmission:fig:backcalc-start-effect}).
This is due to the incorrect assumption that incidence is constant in the past.
Changing this assumption (\eg to exponential growth) modifies the estimated value on the first day, but the rest of the series is insensitive to this change (not shown).
Therefore, the first day's estimate is unreliable, and has been excluded from all other results.
However, the rest of the series is robust to this assumption.
\begin{figure}
    \centering \includegraphics{transmission/backcalc-start-effect}
    \caption[Edge effects in the phenomenological approach]{%
        First seven days of incidence proportion estimate for the whole of England, highlighting the edge effect.
    }
    \label{transmission:fig:backcalc-start-effect}
\end{figure}

@@

In this section, I apply the general mechanistic transmission modelling framework from \cref{E-SEIR} to the CIS prevalence data.
The application of this framework follows \textcite{birrellRealtime}.
The largest modification from the generic framework presented in \cref{E-SEIR} is the parameterization of the WAIFW matrix (\cref{SEIR:sec:WAIFW-application}).
A novel approach (\cref{SEIR:sec:observation}) links this transmission model to the CIS prevalence data; the approach is generalisable to other studies collecting representative prevalence data, \eg REACT.
In the mechanistic modelling literature, this part of the model is referred to as the observation model, however, in the terminology of \cref{E-inc-prev}, it includes both the prevalence process and the observation process.

Having specified the full model (\cref{SEIR:sec:methods-application}), I conducted a simulation study to validate the inference process and identifiability (\cref{SEIR:sec:sim-study}), before applying the model to the CIS data (\cref{SEIR:sec:application}).
In \cref{SEIR:sec:discussion}, I discuss the results; in \cref{transmission:sec:comparison} I compare the results to those in the previous section.

@@

I link the modelled proportion of individuals that are RT-PCR positive to the data using a beta-binomial likelihood (as discussed in \cref{E-biology-data:sec:clustering}).
If on each day $t$, each stratum $a$ is observed to have $y_{at}$ positives out of $n_{at}$ tests, then the likelihood is:
\begin{align}
    \prod_{a,t} \BB (y_{at} \mid n_{at}, \bar{p}_{at}, \rho)
    \label{SEIR:eq:likelihood}
\end{align}
where: $\bar{p}_{at}$ is the mean proportion of individuals in strata $a$ that are RT-PCR positive on day $t$; and $\rho$ is a nuisance parameter controlling the overdispersion of the beta-binomial distribution.
Denoting the proportion of stratum $a$ in state $p_k$ as $p_{ka}(\tau)$, then $\bar{p}_{at} = \int_{t}^{t+1} \sum_{k=1}^l p_{ka}(\tau) d\tau$.

@@

Equating the expressions in \cref{SEIR:eq:fullODEs} and \cref{SEIR:eq:exponential-growth}, gives the following solution for the initial state of most of the compartments.
\begin{align}
    \vec{e}_2(0) &= \vec{i}_1(0) \frac{\psi + 2\gamma}{2\sigma} \\
    \vec{e}_1(0) &= \vec{e}_2(0) \frac{\psi + 2\sigma}{2\sigma} \\
    \vec{p}_0(0) &= \vec{e}_1(0) \frac{\psi + 2\sigma}{\psi + \omega_0} \\
    \vec{p}_1(0) &= \vec{p}_0(0) \frac{\omega_0}{\psi + \omega_1} \\
    \vec{p}_2(0) &= \vec{p}_1(0) \frac{\omega_1}{\psi + \omega_{2}} \\
    \vec{p}_3(0) &= \vec{p}_2(0) \frac{\pi \omega_{2}}{\psi + \omega_{3}}.
\end{align}

@@

I compute the posterior distribution using MCMC.
Specifically, I use an adaptive random-walk Metropolis--Hastings algorithm (discussed in \cref{E-MCMC}).
The proposal distribution is a multivariate normal that adapts to the covariance of the posterior.
The adaptive algorithm is based on \textcite[algorithm 4]{andrieuTutorial}, with the implementation based on \textcite{ghoshApproximate}'s work.
Proposals are accepted or rejected using the standard Metropolis--Hastings acceptance probability.
See \cref{SEIR:MCMC-algorithm} for full details.

@@

I ran a simulation study to investigate the inference procedure's performance and the model's practical identifiability.
I generated 100 simulated datasets representing the East of England using the following procedure.

@@

Next, I solved the ODE system in \cref{SEIR:eq:fullODEs} for each parameter set.
All other aspects of the ODEs, including the data-informed contact matrices, $\matr{\kappa}(t)$, are as previously described in this section.
Population sizes were those of the East of England, and the time period the same as the main analysis.
This means that the simulations created realistic incidence and prevalence curves (see \cref{SEIR:fig:sim-data}).
In particular, the October half-term school holiday is visible, especially in the school-age groups and the November lockdown is visible to some extent in all age groups.
\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/sim/data}}
    \caption[Simulated data]{%
        (A) Incidence and (B) prevalence for the 100 simulated data sets as a proportion of each stratum.
        Prevalence shown is $\bar{p}_{kat}$, the true population prevalence on each day, excluding sampling noise.
        Each line represents one simulation.
    }
    \label{SEIR:fig:sim-data}
\end{figure}

@@

Having generated 100 datasets, I used the inference procedure in \cref{SEIR:sec:inference} to find the posterior distribution conditional on each dataset.
Hence, I ended up with 100 parameter sets distributed according to the priors (the ``generative values''), with an associated dataset and posterior distribution for it.

@@

In this set-up, the parameters $i^+$, $q_c$, and $\psi$ are fully identified; $\pi_0$ and $\rho$ are not identified at all; the other parameters are partially identified, their posterior has moved away from the prior but there is insufficient information in the data to fully overcome the prior (see \cref{SEIR:fig:true-vs-posterior}).
The ability to identify $q_c$ is an advantage of using data which includes measurements on children, unlike data on hospitalisations or deaths.

@@

The \emph{attack rate} is the proportion of the population that have been infected at least once.
In this model it is $1 - \vec{s}$, the proportion susceptible at the time the study ends.
The attack rate is highest in the 12--16 age group in the North West at 68\% (95\% CrI: 66--70\%) (see \cref{SEIR:fig:attack-rates}).
Some of these infections occurred prior to the period covered by this study; during the period of this study 23\% (95\% CrI: 21--24\%) of the population in the North West aged 11--16 was infected.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{SEIR/CIS/attack_rates}}
    \caption[Posterior estimates of the attack rate.]{Posterior median and 95\% CrI of the attack rate in each region and age group.}
    \todo[inline]{Maybe compare this to blood donors, as used in the prior.}
    \label{SEIR:fig:attack-rates}
\end{figure}

@@

These patterns can be interpreted as an instance of the common bias--variance trade-off.
The SEIR model imposes stronger assumptions about the epidemic dynamics.
The assumptions reduce the variance of the estimates, including by making them smoother, but if the assumptions are incorrect, bias is introduced.
Furthermore, the SEIR model does not fully adjust for the non-response bias.
In particular, it makes no adjustment for the under-representation of ethnic minority groups in the CIS.
Adjusting for more covariates would require further stratifying the population and finding an appropriate parameterization of the contact matrix.
This trade-off based on the strength of assumptions has long been known to apply in the backcalculation context~\autocite[e.g.][section 8.3]{brookmeyerBackcalculation}.
The SEIR model can be viewed as a backcalculation method where the transmission part of the model is a smoothing method.
